---
title: "Step 01: Feature Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 01: Feature Selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EBEx)
```

## Introduction
The first step in the EBEx pipeline is to identify potential genes from multiple sources. In this example we show how to obtain the Data-Driven and Disease-Related genes. First, we will define some global variables and relative paths:

```{r parameters, eval = TRUE}
# Parameters
target_var <- "dis_condition"
disease_code <- "C0024117"
# Relative paths
base_dir <- "../test"
raw_dir <- file.path(base_dir, "raw_data")
output_dir <- file.path(base_dir, "feature_selection")
```

To get any feature selection list, we need the train expression matrix and the target variable. Therefore, the first step is to obtain the train and test sets:
```{r preprocessing, eval = TRUE}
# Extract train and test sets
split_data <- obtain_split_data(directory_to_load = raw_dir,
                                file_name = "expression",
                                target_var = target_var,
                                directory_to_save = output_dir)
expression_train <- split_data$train
expression_test <- split_data$test
```

### Example 1: Data-Driven Selection
For obtaining the Data-Driven list we need the Differentially Expressed genes (DEA genes; saved as dea.txt) and the mRMR genes. We have already computed the DEA genes; they should be saved in a dea.txt file with one gene per line. For obtaining the mRMR genes, we need the ./mrmr binary program and the selected threshold (a = 0.031 in this case).

```{r data_driven_example, eval = TRUE}
# This code block requires the mRMR binary
genes_mrmr <- run_feature_selection(
  procedure = "data_driven",
  expression_data = expression_train,
  threshold_value = 0.031,
  target_var = target_var,
  dea_genes = NULL,
  mrmr_genes = NULL,
  directory_to_load = output_dir,
  directory_to_save = output_dir
)
print_message(genes_mrmr)
```

### Example 2: Disease-Related Curated Genes
For obtaining the Disease-Related genes, we use the DisGeNet database. For doing that, we have downloaded the disease-associated genes from curated resources and we can extract genes already associated with a disease code (COPD: C0024117).

```{r disease_related_example, eval = TRUE}
# This code block requires having the DisGeNet database downloaded and knowing the disease code
genes_disgenet <- run_feature_selection(
  procedure = "disease_related",
  target_var = target_var,
  disease_code = disease_code,
  expression_data = expression_train,
  directory_to_load = raw_dir,
  directory_to_save = output_dir
)
print_message(genes_disgenet)
```

